---
title: "Homework 6"
author: "Eunseo Kang"
output:
    html_document:
      toc: true
      toc_float: true
      code_folding: show
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

```{r}
#install.packages("janitor")
library(janitor)
library(tidymodels)
library(tidyverse)
library(glmnet)
library(ggplot2)
tidymodels_prefer()
```
## Tree-Based Models

For this assignment, we will continue working with the file `"pokemon.csv"`, found in `/data`. The file is from Kaggle: <https://www.kaggle.com/abcsds/pokemon>.

The [Pokémon](https://www.pokemon.com/us/) franchise encompasses video games, TV shows, movies, books, and a card game. This data set was drawn from the video game series and contains statistics about 721 Pokémon, or "pocket monsters." In Pokémon games, the user plays as a trainer who collects, trades, and battles Pokémon to (a) collect all the Pokémon and (b) become the champion Pokémon trainer.

Each Pokémon has a [primary type](https://bulbapedia.bulbagarden.net/wiki/Type) (some even have secondary types). Based on their type, a Pokémon is strong against some types, and vulnerable to others. (Think rock, paper, scissors.) A Fire-type Pokémon, for example, is vulnerable to Water-type Pokémon, but strong against Grass-type.


The goal of this assignment is to build a statistical learning model that can predict the **primary type** of a Pokémon based on its generation, legendary status, and six battle statistics.

**Note: Fitting ensemble tree-based models can take a little while to run. Consider running your models outside of the .Rmd, storing the results, and loading them in your .Rmd to minimize time to knit.**

### Exercise 1

Read in the data and set things up as in Homework 5:

- Use `clean_names()`
- Filter out the rarer Pokémon types
- Convert `type_1` and `legendary` to factors

Do an initial split of the data; you can choose the percentage for splitting. Stratify on the outcome variable.

Fold the training set using *v*-fold cross-validation, with `v = 5`. Stratify on the outcome variable.

Set up a recipe to predict `type_1` with `legendary`, `generation`, `sp_atk`, `attack`, `speed`, `defense`, `hp`, and `sp_def`:

- Dummy-code `legendary` and `generation`;
- Center and scale all predictors.

```{r}

data <- read.csv('data/Pokemon.csv') %>% 
  clean_names()
data <- data %>% 
  filter(type_1=="Bug"|type_1=="Fire"|type_1=="Grass"|type_1=="Normal"|type_1=="Water"|type_1=="Psychic") 

data$type_1 <- as.factor(data$type_1)
data$legendary <- as.factor(data$legendary)

set.seed(514222)
Pokemon_split <- initial_split(data, strata = "type_1", prop=0.7)

Pokemon_train <- training(Pokemon_split)
Pokemon_test <- testing(Pokemon_split)

Pokemon_fold <- vfold_cv(Pokemon_train, v = 5, strata= "type_1")

Pokemon_recipe <- 
  recipe(formula = type_1 ~ legendary+generation+sp_atk+attack+speed+defense+hp+sp_def, data = Pokemon_train) %>% 
  step_novel(all_nominal_predictors()) %>% 
  step_dummy(legendary,generation) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors())
```
### Exercise 2

Create a correlation matrix of the training set, using the `corrplot` package. *Note: You can choose how to handle the continuous variables for this plot; justify your decision(s).*

What relationships, if any, do you notice? Do these relationships make sense to you?

```{r}
#install.packages("corrplot")
library(corrplot)
cor(Pokemon_train[sapply(Pokemon_train, is.numeric)]) %>% 
  heatmap()
```

***I find very strong positive relationship with the ID number of the pokemon and the generation. It is almost 1 since it is natural that the same pokemon should be in the same generation. Total shows also generally positive relationship with each stat, and it makes sense since total is the sum of all stats.***

### Exercise 3

First, set up a decision tree model and workflow. Tune the `cost_complexity` hyperparameter. Use the same levels we used in Lab 7 -- that is, `range = c(-3, -1)`. Specify that the metric we want to optimize is `roc_auc`. 

Print an `autoplot()` of the results. What do you observe? Does a single decision tree perform better with a smaller or larger complexity penalty?

```{r}
library(rpart.plot)
tree_spec <- decision_tree() %>%
  set_engine("rpart")
class_tree_spec <- tree_spec %>%
  set_mode("classification")
class_tree_wf <- workflow() %>%
  add_model(class_tree_spec %>% set_args(cost_complexity = tune())) %>%
  add_recipe(Pokemon_recipe)

param_grid <- grid_regular(cost_complexity(range = c(-3, -1)), levels = 10)
```

```{r, eval=F}
tune_res <- tune_grid(
  class_tree_wf, 
  resamples = Pokemon_fold, 
  grid = param_grid, 
  metrics = metric_set(roc_auc)
)
save(tune_res, file = "hw6.rda")

```

```{r}
load(file = "hw6.rda")
best_complexity <- select_best(tune_res, metric="roc_auc")

class_tree_final <- finalize_workflow(class_tree_wf, best_complexity)

class_tree_final_fit <- fit(class_tree_final, data = Pokemon_train)

autoplot(tune_res)
```

***It performs better with smaller complexity penalty. ***

### Exercise 4

What is the `roc_auc` of your best-performing pruned decision tree on the folds? *Hint: Use `collect_metrics()` and `arrange()`.*

```{r}
collect_metrics(tune_res) %>% 
  arrange()
```


***It is 0.6555111. ***

### Exercise 5

Using `rpart.plot`, fit and visualize your best-performing pruned decision tree with the *training* set.

```{r}
class_tree_final_fit %>%
  extract_fit_engine() %>%
  rpart.plot()
```


### Exercise 5

Now set up a random forest model and workflow. Use the `ranger` engine and set `importance = "impurity"`. Tune `mtry`, `trees`, and `min_n`. Using the documentation for `rand_forest()`, explain in your own words what each of these hyperparameters represent.

Create a regular grid with 8 levels each. You can choose plausible ranges for each hyperparameter. Note that `mtry` should not be smaller than 1 or larger than 8. **Explain why not. What type of model would `mtry = 8` represent?**

```{r}
rf_spec <- rand_forest(mtry = tune(), trees=tune(), min_n=tune()) %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("classification")
rf_wf <- workflow() %>%
  add_model(rf_spec) %>%
  add_recipe(Pokemon_recipe)
```


***mtry means the number of predictors that will be randomly sampled at each split when creating the tree models. `trees` means the number of trees contained in the ensemble. `min_n` means the minimum number of data points in a node that are required for the node to be split further.***

```{r}
grid <- grid_regular(mtry(range = c(1, 8)), trees(range = c(0, 10)), min_n(range=c(0,10)), levels = c( mtry=8, trees=8, min_n=8))
```

### Exercise 6

Specify `roc_auc` as a metric. Tune the model and print an `autoplot()` of the results. What do you observe? What values of the hyperparameters seem to yield the best performance?

```{r, eval=F}
#install.packages("ranger")
tune <- tune_grid(
  rf_wf, 
  resamples = Pokemon_fold, 
  grid = grid, 
  metrics = metric_set(roc_auc)
)
save(tune, file = "hw6_2.rda")
```

```{r}
load(file="hw6_2.rda")
autoplot(tune)
```

### Exercise 7

What is the `roc_auc` of your best-performing random forest model on the folds? *Hint: Use `collect_metrics()` and `arrange()`.*

### Exercise 8

Create a variable importance plot, using `vip()`, with your best-performing random forest model fit on the *training* set.

Which variables were most useful? Which were least useful? Are these results what you expected, or not?

### Exercise 9

Finally, set up a boosted tree model and workflow. Use the `xgboost` engine. Tune `trees`. Create a regular grid with 10 levels; let `trees` range from 10 to 2000. Specify `roc_auc` and again print an `autoplot()` of the results. 

What do you observe?

What is the `roc_auc` of your best-performing boosted tree model on the folds? *Hint: Use `collect_metrics()` and `arrange()`.*

### Exercise 10

Display a table of the three ROC AUC values for your best-performing pruned tree, random forest, and boosted tree models. Which performed best on the folds? Select the best of the three and use `select_best()`, `finalize_workflow()`, and `fit()` to fit it to the *testing* set. 

Print the AUC value of your best-performing model on the testing set. Print the ROC curves. Finally, create and visualize a confusion matrix heat map.

Which classes was your model most accurate at predicting? Which was it worst at?

## For 231 Students

### Exercise 11

Using the `abalone.txt` data from previous assignments, fit and tune a random forest model to predict `age`. Use stratified cross-validation and select ranges for `mtry`, `min_n`, and `trees`. Present your results. What was the model's RMSE on your testing set?